Background for transient faults in asynchronous circuits

### 2019-07-04 draft_together.tex
before major revision

\begin{document}
\chapter{Introduction}
Digital circuits is the basis of many of present day electronics.  Digital circuits has signals which are represented as discrete values.  This is achieved by choosing a range of voltages that correspond to each discrete value.  Many digital circuits use binary values, that is the signals can only take on the value `logic 0' or `logic 1' (or low/high).  One advantage of only allowing discrete values is that digital circuits are much more robust to noise fluctuations in voltage.  Disregarding the lower level implementation, one can also abstract the function of digital circuits in terms of gates.\\

@@

\section{Organization of this work}
%why are async circuits important, timing delay etc
Because these various faults can interfere with the function of the circuit and even result in harmful operations (catastrophic failures), there has been a lot of work put into making digital circuits fault tolerant. Most of the fault tolerant solutions has been investigated for synchronous digital circuits.  Because the timing difference between different signals arriving do not have to be considered for synchronous circuits, they can use error correcting codes and redundancy techniques such as triple modular redundancy (TMR). Unfortunately many of these techniques have problems if directly implemented in asynchronous circuits.\\

In this work, we propose a new fault tolerant circuit transformation on asynchronous semi-modular control circuits.  We also propose a method of comparing the fault tolerant circuit with the original circuit, by comparing traces with the half circuit.  The fault free half of the fault tolerant circuit behaves like the original circuit under single transient faults.  In addition, we assume a gate level model of the circuit and allow the transient to occur on any of the wires thus solving the transient fault problem for a bigger class of circuits than the current state-of-the-art.\\

In the rest of this work, in chapter two, I will first introduce the different types of common faults and their causes, as well as the common fault models that researchers use when investigating these faults.  Then I will go into depth about the asynchronous circuit models that are relevant to this work.  I will then go through the existing body of work on fault tolerant solutions for both types of circuits with an emphasis on asynchronous circuits.  In chapter three, I will show why the existing fault tolerant methods are insufficient and propose a new fault free method for asynchronous transient faults.  I will prove using logical proofs as well as formal verification methods to show why my new method always works.  In the final chapter, chapter four, I will conclude on how to better build fault free asynchronous circuits and suggest some possible new research directions.
%Overview:  why is the problem important  \\  maybe I can extend the Introduction a bit more
\chapter{Background}
\section{Faults}
Faults can be split into permanent and non-permanent faults.  Non-permanent faults are further split into transient and intermittent faults \cite{jha_gupta_2003}. %copied 
A fault is defined as the physical difference between the good or correct system and the current system.  An error may be caused by a fault, which is when the state of the system differs from the state of the good system and thus unable to perform some specified function.  Hidden faults are also possible where the fault does not cause an error \cite{jha_gupta_2003}. \\

@@

The approach that we use to make circuits fault tolerant will be described in the next chaper.  Some of the previous works makes additional assumptions such as the checking circuit is fault free or that only one fault can occur on an arbitrarily large transistor gate.  We allow for standard two-input gates and assume that all gates in our model are subject to faults.  Our fault tolerant transformation procedure is for a class of circuits, semi-modular control circuits, which includes many circuits.

\chapter{Fault Tolerant Circuit Under Normal Operation and Single Transient Fault}
\chaptermark{Fault Tolerant Circuit}
\section{Traces and circuit equivalence}
We solve the problem of single transient faults by taking a semi-modular circuit (the original circuit) and applying a transformation to the circuit by duplicating all its circuit elements.  Due to the asynchronous nature of how signals are processed, corresponding wires in the resulting half circuits may have different values.  To solve this problem we add C-elements to synchronize the two versions of the same signal.   An example of this transformation applied to the simple inverter circuit is shown in figure \ref{fig:dupschemeex}.  Both the duplication and the additional C-elements configuration are shown.  In the rest of this chapter we define a way to compare two circuits and to show their equivalence.  Then we propose our transformation method and show that under normal operation (no faults), the fault tolerant circuit is equivalent to the original.  Finally we show that if a single transient fault occurs, one side of the fault tolerant circuit is equivalent to the original and we can still extract the signals that we desire.\\
\begin{figure}
  \centering
    \includegraphics[width=.9\textwidth]{transf3Csimp}
  \caption{a) original circuit b) after the fault tolerant transformation}
  \label{fig:dupschemeex}
\end{figure}

Other than the basic terminology for asynchronous circuits, we also want terminology to compare two circuits based on their function.  
We first define a {\em trace}, to describe a path the circuit can take as it follows the state graph.  Different paths can occur because there may be multiple transitions from each state.  We call the succession of states and transitions the trace which is defined as follows:
\begin{definition}A {\em trace} $\sigma$ of a circuit is a sequence of states of finite or infinite length and $s_i$ is the $i^{th}$ term of the sequence.  %Formally this is written as $\sigma: 0,\ldots ,n-1 \to S$ and $n \in \mathbb{N}$.
 Also every pair of consecutive states in a trace are transitions.  % such that $T(s_i,s_{i+1})$ for $ 0\leq i\leq n-2$
We also use the $\sigma_n$ as shorthand to denote the sequence of the first n states of $\sigma$\end{definition} 
Thus a circuit has a set of traces.  A circuit also has a set of initial states where the traces can begin in.  The set of initial states are the same as the set of states in the state graph.\\%new 4/30

Note that we allow the circuit to have external wire inputs and stopping states.  
\begin{definition}A {\em stopping state} is a state where there are no excitations in the wires in the circuit and the circuit is waiting for the external inputs.\end{definition} 
A trace of finite length ends in a stopping state where the external inputs never arrive and the circuit is stopped.  A related concept is a deadlocked state which is a state where the circuit prematurely stops and defined as follows:
\begin{definition}A {\em deadlocked state} is a state where there are no excitations in the wires in the circuit and is not a stopping state.\end{definition}
A deadlocked state does not appear in a well designed circuit, but may appear after a circuit transformation (i.e., the new circuit stops at states where the original circuit do not), or when there are faults.  Thus we want to design a circuit transformation where there are no deadlocked states.\\ %end new
%Note that our definition of trace also includes the case where there is only one transition from each state of a state graph, which results in a linear state graph.

\begin{figure}
  \centering
    \includegraphics[width=.6\textwidth]{sm_counter2}
  \caption{a) original circuit for wire w b) an erroneous initialization}
  \label{fig:sm_counter2}
\end{figure}

The original circuit is semi-modular by design.  But the design only covers cases where the wires are in a state in the state graph.  That is, hazards may occur if the circuit goes into an unreachable state of the state graph and the circuit is no longer semi-modular and furthermore becomes unpredictable.  This does not happen in the fault-free case where the circuit is initialized to a state in the state graph and the wires inside the logic blocks are properly initialized.  However when faults do occur, the circuit may become unpredictable which is a reason why we need to transform the circuit to protect against faults.  An example of what might happen when the circuit is not in a semi-modular state is shown in figure \ref{fig:sm_counter2}.  The logic circuit for wire $w$ is taken out of a semi-modular circuit.  However, if the state of the $w,x,y,z$ wires are $(0,1,0,1)$ and the internal wires after the AND and OR gate are both initialized to 1, this is not semi-modularity consistent.  The output wire $w$ is excited to 1 while the inputs from $x,y,z$ will cause $w$ to transition to 0.  Depending on the timing of the gates, $w$ may stay at 0 or experience a glitch to 1 then back to 0.  The extra layers of gates in the logic block can be thought of as extra delays that can store and propagate the wrong values.  More detailed discussion on how to initialize and keep the circuit semi-modular is explored in the subsequent section.\\

Given two circuits, how can we determine when their basic functionality is the same, especially if they are implemented with different gates and circuit elements?  We define a notion of equivalence between two circuits through their sets of traces.  
%We want the underlying state graph representation of the circuits to be equivalent.  However only the signals and the sequence of transitions from the circuit is observable.  In other words the trace is observable.  Thus we define circuits as equivalent if the set of traces of the two circuits are the same. [We use the assumption that if the set of traces are equal then the state graph is equal.]
\begin{definition} Two circuits A and B are {\em equivalent} if the set of all traces from circuit A is equal to circuit B.  In other words, $\{\sigma^A |\sigma^A$ is a possible trace in $A\}=\{\sigma^B |\sigma^B$ is a possible trace in $B\}$.
\end{definition}

We are now ready to explore how to make a circuit fault-tolerant.
%good enough for now, do I need if and only if? <-NO, xxx if xxx in math lingo is iff

%An additional property of the asynchronous circuit is that one can view each wire as a separate black box.  For a wire $v$, inputs from the rest of the circuit ($I_v$ and $I_v \subseteq$ V) feed into a gate and outputs the wire value $v$.  We call the instantaneous value of the gate for $v$ of a state $s$ as $f_v(s|_I)$, where f is the function $f:[I \to \{0,1\}] \to \{0,1\}$ and $s|_I$ is the state $s$ projected on the input variables $I_v$.  If inputs to the gate remain constant, the output $v$ will eventually take the value specified by $f_v(s|_I)$  %maybe change v again, also the f(s) vs f(I) thing...
\section{Fault Tolerant transformation}

We can make an existing semi-modular circuit fault-tolerant by the following transformation as shown in figure \ref{fig:dupscheme}: make a duplicate copy of the logic blocks of each wire $w$ and label the duplicate wires $w^A$ and $w^B$.  We also label the duplicated logic blocks as belonging to circuit A or circuit B in agreement to the wire naming.  Then we connect the output of the logic blocks to three layers of C-elements as shown in figure \ref{fig:dupscheme}.  We label the intermediate wires as $c1^A$, $c2^A$, $c3^A$ for circuit A and $c1^B$, $c2^B$, $c3^B$ for circuit B.  In the first layer of C-elements, $c1^A$ and $c1^B$ are outputs of the logic blocks and inputs to two C-elements outputting $c2^A$ and $c2^B$.  These are then the inputs to the second layer of C-elements outputting $c3^A$ and $c3^B$.  These are then the inputs to the third layer of C-elements outputting $w^A$ and $w^B$.  We can label one of the final C-element outputs as $w^A$ and the other output as $w^B$.  These then connect to other logic blocks in circuit A and circuit B respectively.  Note that when labelling the C-elements and its outputs, which side is A and B can be somewhat arbitrary because the output pairs are symmetric.  The symmetry is broken when connecting the final output wire $w^A$ to all the circuit A gate inputs and wire $w^B$ to all the circuit B gate inputs.  All circuit elements with the A label is part of the A-half circuit and all circuit elements with the B label is part of the B-half circuit.\\%(Note that the outputs of the C-elements can be arbitrarily named and connected to the next gates, and the next gates can also be arbitrarily named $A$ or $B$ as long as it does not create any inconsistencies in the naming convention, ie. all $A$ wires feed into the $A$ gates)  %the only wires that really matter are the wires that have isochronic forks, these must input to same A or B, everything else is just a naming difference
\begin{figure}
  \centering
    \includegraphics[width=\textwidth]{circuitforproof3}
  \caption{Fault tolerant transformation: duplicate the original logic for each output wire and adding 3 layers of C-elements}
  \label{fig:dupscheme}
\end{figure}

We want to show that a half circuit from the fault tolerant circuit has correct function.  To do this, we define a correspondence between states of the original circuit and states of the fault tolerant circuit.  As a result of the circuit transformation, a state of the fault tolerant circuit has bits for 8 times as many wires as a state in the original circuit.  Half of those bits are obviously from the A half and half are from the B half of the circuit.  The state graph for a fault tolerant circuit is defined on this combined state.  We define the half-circuit mapping $h^A: s^{\mathit{full}} \to s^A$ where $s^{\mathit{full}}$ is the combined state of a fault tolerant circuit and $s^A$ consists of wires $w^A$ that correspond to the original wires $w$ in the original duplication procedure.  The half-circuit mapping $h^B$ is similarly defined to map $s^{\mathit{full}}$ to $s^B$.  The added wires such as $c1^A$, $c1^B$, $c2^A$, $c2^B$, $c3^A$, $c3^B$ from the duplication procedure are hidden by $h^A$ or $h^B$.  We extend $h^A$ to traces by using the following two steps: first, $h^A$ is applied individually to each state in a trace $\sigma$, and likewise for $h^B$.  %not sure if I want the mapping to be from variable to variable or specific state to state
%For a trace in the fault tolerant circuit we project the full states onto a shortened list of wires (ie the wires corresponding to the original circuit). 
The resulting trace may have successive repeated states.  So next we delete all but one of the neighboring repeats to obtain a valid trace on the shortened list of wires.  The result should be a trace from the original circuit.  
Using this mapping we will show that the original circuit is equivalent to half of the fault tolerant circuit in both the absence and presence of a single transient fault.
%do I need to define all the new transitions in duplicated circuit?  maybe
\\

%[Why is this the best we can do?]
We compare each half circuit because it is not easy to synchronize and compare the signals between both halves, especially if we reuse the logic blocks from the original circuit.  A simple idea we tried was to insert another C-element before the input is fed into the logic, to synchronize the wires on both halves.  The problem with that is in the original circuit, a wire value change propagates to all inputs at the same time.  Adding another layer of C-elements creates delays at the input to logic blocks and the isochronic fork assumption is violated.  Thus the resulting circuit will not function according to the specified state graph.  (an example can be provided but not sure if needed?)\\  
\begin{figure}
  \centering
    \includegraphics[width=.8\textwidth]{counterexhalf}
  \caption{State graph example to show how b can transition before a or c finishes in both halves in the fault tolerant circuit}
  \label{fig:counterexhalf}
\end{figure}

Next we examine the traces of each half circuit.  Figure \ref{fig:counterexhalf} shows a subsection of a semi-modular state graph where each state is given a nominal name (state 0, state 1, ... etc) and each transition is labeled with the transitioning wire.  Suppose we apply our fault tolerant transformation to such a circuit based on this state graph.  Both halves in the fault tolerant circuit will transition based on this state graph, with the added caveat that for a wire to transition at some state, it must be excited in both halves.  Starting in {\em state 0}, it is possible for the A-half to transition to {\em state 1} ($a^A$ transitions) and for the B-half to transition to {\em state 2} ($c^A$ transitions).  We assume that $a^B$ and $c^A$ do not complete their transitions yet.  Immediately, in the A-half, $b^A$ is excited and in the B-half, $b^B$ is excited.  Then, suppose the transition on wire $b^A$ and $b^B$ occurs.  That is, the A-half transitions to {\em state 3} and the B-half transitions to {\em state 5}.  And finally, the excited $a^B$ and $c^A$ transitions so that both halves are in {\em state 6}.  The trace so far for the A-half is (state 0, state 1, state 3, state 6) while for the B-half it is (state 0, state 2, state 5, state 6).  Thus the two halves need not have the same trace.  One criticism may be that the current fault tolerant transformation synchronizes at the outputs of the logic blocks.  If we want to reuse the same logic blocks but synchronize the signals earlier, such as replacing the final C-element in a standard C implementation with a 4C-element, the issue of different traces in the two halves will still occur.  We have already shown earlier that synchronizing the input signals while keeping the original logic block can also result in error.  This points to looking at properties of each half circuit as the right thing to do.\\

Another wrong approach for the fault tolerant transformation is that if we take the following view for transitions, a transition of wire $w$ in the original circuit is complete if both $w^A$ and $w^B$ transitions in the fault tolerant circuit, then some signals may transition out of order.  We use the earlier example of figure \ref{fig:counterexhalf} to show this as well.  Following the same transitions outlined above, $b^A$ and $b^B$ can finish their transitions before $a^B$ and $c^A$ transitions.  Then the $b+$ transition is actually completed before $a+$ or $c+$ and the combined trace no longer maps back to a valid trace in the original state graph.  Another approach we may take is to record the first transition of a wire in either half of the fault tolerant circuit.  For example, if $w^A=w^B$ and then $w^A$ transitions, we map this to a transition on $w$ in the original circuit.  This approach ensures the resulting trace is a trace in the original circuit (can be shown by Lemma 1 and Lemma 2.1 in the next section), however, in the presence of faults, it is easy to see that it does not provide any fault tolerance.\\

For the aforementioned reasons, we choose to use the $h^A$ and $h^B$ mappings and compare each half circuit against the original circuit.  Out of the various comparisons methods and circuit transformations that we tried, it is the most reasonable and results in some nice properties.  We will next demonstrate that the half circuits are equivalent to the original circuit in the absence of faults.  Then we show that a similar property is true even when a single fault may occur.

\section{Normal operation analysis}
Now we will discuss the proper wire initializations before going into the Lemmas.  In normal operation, the fault tolerant circuit must have wire values consistent with semi-modular behavior in the original circuit.  
This includes the internal wires which exist in the logic blocks in the original circuit and through duplication also exist in the fault tolerant circuit.  The actual values of these wires do not matter if the inputs to the logic blocks follow the semi-modularity property, and if the internal wires are initialized properly.  This arises from the original synthesis method to synthesize the circuit from an underlying state graph.  
Additionally, it does not suffice for the inputs to the logic blocks to be from a valid trace of the state graph; the timing matters as well.  The semi-modularity property must be preserved:  if the logic block output is excited then it must remain excited until a transition on its output occurs.  Since certain inputs which are valid states and arise from valid transitions may cause the logic block to become unexcited before a transition occurs and thus break the semi-modularity property, we require this more stringent criteria.  %maybe needs an example
Thus to satisfy semi-modularity, the inputs wait at certain input states until the output transitions.  Then the internal wires of the logic block are termed {\em semi-modularity consistent} if the following properties are true:  if the inputs follow semi-modularity then there are no glitches at the output; and once the output is excited it remains excited until the output transitions. \\%newly added

What values to initialize the wires of a circuit to is very important.  We want certain properties to hold for all states that the circuit goes through, and such properties must also hold for the initial state.  On one hand, the initial state $s$ must be a state in the underlying state graph.  Thus the wire values is initialized to be consistent with that property.  On the other hand, as mentioned previously, the internal wires at the initial state (and each state thereafter) must be semi-modularity consistent.  We assume that we can initialize the wire values of the original circuit to satisfy these two properties.  Given this, we pick convenient initializations for the fault tolerant circuit.  The following procedure covers how to initialize both the wires from the duplication as well as the extra $c1^A$, $c1^B$, $c2^A$, $c2^B$, $c3^A$, $c3^B$ wires that are introduced.
%For all initial states we assume that we know the values of every wire in the circuit, intermediate and visible wires.  The visible wire values are the output wires in the circuit $w$, and the wires $c1$, $c2$ and $c3$'s. The intermediate wires are all other wires, which consists of all the wires within the logic blocks for each output wire.  Given these we will show how to define a corresponding initial state in the duplicated circuit.  %should I delete intermediate?
\begin{definition} A {\em valid initial state} for the A half of the fault tolerant circuit is constructed from an initial state of the original circuit.  The wire values in the logic block and $w$ in the original circuit are copied to the corresponding wires in the A half.  The additional wires from the C-element chain are also assigned the same value as $w$.  That is, first assign to $w^A$ the value of $w$, then assign to $c1^A$ $c2^A$ $c3^A$ the value of $w^A$ as well. The same construction applies for the B half circuit as well.
\end{definition}
%\textbf{
This definition is stronger than it needs to be but is convenient.\\

%To avoid overly clunky sentences, we will assume that fault tolerant circuits start from the same valid intial state in both halves.  One exception to this assumption is Lemma 1 where only one half circuit is needed to start from a valid initial state since the property is true regardless of the wire values of the other half.\\ 
%For all subsequent discussion in this document, when we talk about the trace of a fault tolerant circuit we assume that both half circuits start from the same valid initial state.  I will not include this assumption in the following lemma definitions to avoid complicated wording.  %}  
%The one exception is for Lemma 1 where we only require one half circuit to start from a valid initial state.  \\

To prove claims about the fault tolerance of this circuit, we first prove that the circuit behaves similarly to the original circuit under normal operation.  That is, the set of traces of the fault tolerant circuit after applying a mapping of $h^A$ is equivalent to that of the original circuit.  And likewise after applying a mapping of $h^B$ to the fault tolerant circuit.  We show this through several Lemmas. Lemma 1 shows for each fault-free half of a circuit, if a transition occurs, it must be a transition in the state graph of the original circuit.  This is a strong result; it is true for arbitrary behavior of
signals in the other half circuit, so long as those signals are well formed Boolean values and also does not cause the fault-free half to deviate from well formed Boolean values.  
This result is also significant later in proving fault tolerance under a single transient fault since one of the halves is fault-free.  %as long as it does not violate digital abstraction
%One additional assumption that we make about the gates and wire values (including in the presence of faults) for Lemma 1 to hold is that wire values are strictly 0 or 1, and transitions from excited gates also only produce 0 or 1 and no in between values, no matter the duration of signals from the input wires.
\\

Note that Lemma 1 allows for the possiblity of deadlock.  Lemma 2 shows deadlock does not occur during normal operation.  Combining these two Lemmas, we can show that all traces in the duplicated circuit can be mapped to a trace in the original circuit.\\
% and all traces in the original can be mapped to duplicated after h mapping?  <-this is obvious though by setting the A half and B half to both transition according to original trace
%[so no deadlock in normal operation??]

This leads to the definition and proof of Lemma 1 as follows:

\begin{lemma}[1]
If $\sigma$ is the trace of the fault tolerant circuit where the A half begins in a valid initial state and is fault free, then the following three properties are true:
\begin{itemize}
\item P1: The half-circuit mapping of a prefix of $\sigma$, $h^{A}(\sigma_n)$, is also a prefix of a trace of the original circuit, regardless of the wire values in the B half.
\item P2: The logic blocks in the A half are semi-modularity consistent for all states in $\sigma$
\item P3: For all states $s$ in $\sigma$, a wire $w$ in state $h^A(s)$ is excited in the original circuit if and only if exactly one of the following is true in the fault tolerant circuit: $c1^A$ is excited or $c1^{A}\neq c2^{A}$ or $c2^{A}\neq c3^A$ or $c3^{A}\neq w^A$.  We also call this condition: wireblock $w$ is excited.  (No more than one of the four conditions mentioned may be true regardless if $w$ is excited or unexcited.)  A wire $w$ in state $h^A(s)$ is unexcited in the original circuit if and only if none of the four conditions are true in the fault tolerant circuit.
\end{itemize}

%Let $\sigma_n$ be the sequence of the first $n$ states of a trace $\sigma$ of the fault tolerant circuit (the trace is possibly finite).

@@

%If the duplicated circuit is in state $s_n$ and the original circuit is in state $h^A(s_n)$, and for all wires $w$ that are excited in this state $f_w(s_n|_I)\neq w$ in original circuit, then $f_w(s_n|_I) \neq x^A$ or $x^A \neq w^A$ in duplicated circuit (exactly one of these is true).  
%In addition, if wire $w$ is not excited in the original circuit $f_w(s_n|_I)=w$, then in the duplicated circuit half $f_w(s_n|_I)=x^A=w^A$.  
\end{lemma}

>>>>

\begin{document}
\chapter{Introduction}
Digital circuits and its widespread use has spurred on many of present day technology.  Digital circuits differs from analog circuits in that its signals (voltages) can be represented as discrete values whereas signals in analog circuits are represented by continuous voltages.  This is achieved by delimiting the ranges of voltages that correspond to a particular discrete value.  Many digital circuits use binary values, that is the signals can only take on the value `logic 0' or `logic 1' (or low/high).  One advantage of only allowing discrete values is that digital circuits are much more robust to noise fluctuations in voltage.  Disregarding the lower level implementation, one can also abstract the function of digital circuits in terms of gates.\\

@@

\section{Organization of this work}
%why are async circuits important, timing delay etc
Because these various faults can interfere with the function of the circuit and even result in harmful operations(catastrophic failures), there has been a lot of work put into making digital circuits fault tolerant. Most of the fault tolerant solutions has been investigated for synchronous digital circuits.  Because the timing difference between different signals arriving do not have to be considered for synchronous circuits, they can use error correcting codes and redundancy techniques such as triple modular redundancy (TMR). Unfortunately many of these techniques have problems if directly implemented in asynchronous circuits.\\

In the rest of this work, in chapter two, I will first introduce the different types of common faults and their causes.  As well as the common fault models that researchers use when investigating these faults.  Then I will go into depth about the asynchronous circuit models that are relevant to this work.  I will then go through the existing body of work on fault tolerant solutions for both types of circuits with an emphasis on asynchronous circuits.  In chapter three, I will show why the existing fault tolerant methods are insufficient and propose a new fault free method for asynchronous transient faults.  I will prove using logical proofs as well as formal verification methods to show why my new method always works.  In the final chapter, chapter four, I will conclude on how to better build fault free asynchronous circuits and suggest some possible new research directions.
%Overview:  why is the problem important  \\  maybe I can extend the Introduction a bit more
\chapter{Background}
\section{Faults}
Faults can be split into permanent and non-permanent faults.  Non-permanent faults are further split into transient and intermittent faults \cite{jha_gupta_2003}. %copied 
A fault is defined as the physical difference between the good or correct system and the current system.  An error may be caused by a fault, which is when the state of the system differs from the state of the good system and thus unable to perform some specified function.  Hidden faults are also possible where the fault does not cause an error \cite{jha_gupta_2003}. \\

@@

In some of the previous work they make additional assumptions such as the checking circuit is fault free.  We propose an asynchronous circuit transformation on a more general type of circuits, control circuits.  The transformed circuit can function under single transient faults.  In addition, we assume a gate level model of the circuit and allow the transient to occur on any of the wires thus solving the transient fault problem for a bigger class of circuits.

\chapter{Fault Tolerant Circuit Under Normal Operation and Single Transient Fault}
\section{Normal operation analysis}
As mentioned in the previous section on Speed Independent circuits, circuits can be modeled by a state graph.  
As the circuit follows the state graph, there may be multiple transitions per state and each path through the states and transitions may be different.  We want to record the succession of states and transitions and so we define a trace as follows:
\begin{definition}A {\em trace} $\sigma$ of a circuit is a sequence of states of infinite length where $\sigma_i$ is the $i^{th}$ term of the sequence.  $\sigma: \mathbb{N} \to S$ and $T(\sigma_i,\sigma_{i+1})$ for $\forall i \in \mathbb{N}$.\end{definition} 
%add finite traces
% all states or just the state graph ones

We want to define a notion of equivalence between two cicrcuits.  What we observe are the succession of signals which form the trace.  Thus we define circuits as equivalent if the traces are the same. 
\begin{definition} Two circuits A and B are {\em equivalent} if the set of all traces from circuit A is equal to circuit B $\{\sigma^A |\sigma^A$ is a possible trace in $A\}=\{\sigma^B |\sigma^B$ is a possible trace in $B\}$.
\end{definition}
%good enough for now, do I need if and only if?

%An additional property of the asynchronous circuit is that one can view each wire as a separate black box.  For a wire $v$, inputs from the rest of the circuit ($I_v$ and $I_v \subseteq$ V) feed into a gate and outputs the wire value $v$.  We call the instantaneous value of the gate for $v$ of a state $s$ as $f_v(s|_I)$, where f is the function $f:[I \to \{0,1\}] \to \{0,1\}$ and $s|_I$ is the state $s$ projected on the input variables $I_v$.  If inputs to the gate remain constant, the output $v$ will eventually take the value specified by $f_v(s|_I)$  %maybe change v again, also the f(s) vs f(I) thing...

We make a fault tolerant circuit by the following circuit transformation: make a duplicate copy of the logic blocks to each wire.  We label the duplicated circuit halves as circuit A and circuit B.  Then for each output wire $w$ we connect the output of the logic blocks from each circuit half to three layers of c-elements as shown in figure \ref{fig:dupscheme}.  We label the intermediate wires as $c1^A$, $c2^A$, $c3^A$ and $c1^B$, $c2^B$, $c3^B$ for circuits A and B respectively.  In the first layer of C-elements, $c1^A$ and $c1^B$ are outputs of the logic blocks and are inputs to two C-elements producing $c2^A$ and $c2^B$.  These are then the inputs to the second layer of C-elements producing $c3^A$ and $c3^B$.  These are then the inputs to the third layer of C-elements producing $w^A$ and $w^B$.  We can (arbitrarily) label one of the final C-element outputs as $w^A$ and the other output as $w^B$.  These then connect to other gates in circuit A and circuit B respectively.  (Note that the outputs of the C-elements can be arbitrarily named and connected to the next gates, and the next gates can also be arbitrarily named $A$ or $B$ as long as it does not create any inconsistencies in the naming convention, ie. all $A$ wires feed into the $A$ gates)  %the only wires that really matter are the wires that have isochronic forks, these must input to same A or B, everything else is just a naming difference
\begin{figure}
  \centering
    \includegraphics[width=\textwidth]{circuitforproof3}
  \caption{Duplicating the original logic for each output wire and adding 3 layers of C-elements}
  \label{fig:dupscheme}
\end{figure}

Because the transformed circuit has some extra wires, we want to define a correspondence to be able to compare between the wires of the original circuit and that of the duplicated circuit halves.  For the duplicated circuit, a combined state is the pairing of the state of circuit A and the state of circuit B. It has eight times the number of wires for each output wire in the original circuit.  The state graph for a duplicated circuit is defined on this combined state.  We define the half-circuit mapping $h^A: s^{full} \to s^A$ where $s^{full}$ is the combined state of a duplicated circuit and $s^A$ are made up of the wires $w^A$ that correspond to the original wires $w$.  The half-circuit mapping $h^B$ is similarly defined to map to $s^B$.  Note that the $h(\sigma)$ mappings are overloaded for a trace, where it outputs a mapped state for each of the states in a trace.  %not sure if I want the mapping to be from variable to variable or specific state to state
For a trace in the duplicated circuit we project the full states onto a shortened list of wires (ie the wires corresponding to the original circuit). The resulting trace may have successive repeated states and we delete the repeats to obtain a valid trace on the shortened wires.  %do I need to define all the new transitions in duplicated circuit?  maybe
\\

We want to show this circuit behaves similar to the original circuit under normal operation.  In other words we want to prove that after applying either mapping of $h^A$ or $h^B$, the resultant circuit and the original circuit are equivalent.  We first prove lemma 1 which states for each half of a circuit, if a transition occurs it must follow the transitions of the state graphs in the original circuit.  However this does not guarantee that such a transition always exists so we also need to prove lemma 2 to show that a transition always exists and the circuit cannot deadlock.  These two properties combined together show that all traces in the duplicated circuit can be mapped to a trace in the original circuit.\\

Note that there are some extra wires in the logic blocks in both the original and duplicated circuit.  We actually don't worry about the exact values of these wires other than that if the inputs follow Semi-Modularity then the logic block as an entirety behaves with Semi-Modularity (as long as it is initialized appropriately).  Here it does not suffice for the inputs to be a sequence of valid states and state transitions in the state graph.  This is because the Semi-Modularity property of the logic block is that if the logic block output is excited then it must remain excited until a transition occurs.  Since certain inputs, even if they are valid states, may cause the logic block to become unexcited before a transition occurs, we want a more stringent criteria.  The inputs must uphold the Semi-Modularity property which waits at certain input states until the output transitions, in addition to being valid states and state transitions from the state graph.  Then if the wires within the logic block is consistent with Semi-Modularity, that is if the inputs follow Semi-Modularity then there are no glitches at the output and if the output is excited it remains excited until the output changes. \\%newly added

The initial state in the original circuit must be assigned wire values that is consistent with Semi-Modularity.  For all initial states we assume that we know the values of every wire in the circuit, intermediate and visible wires.  The visible wire values are the output wires in the circuit $w$, and the wires $c1$, $c2$ and $c3$'s. The intermediate wires are all other wires, which consists of all the wires within the logic blocks for each output wire.  Given these we will show how to define a corresponding initial state in the duplicated circuit.  %should I delete intermediate?
\begin{definition} A {\em valid initial state} for the A half of a duplicated circuit is defined as follows.  Given any initial state in the original circuit $s \in S$, assume that for each wire $w$, the wire values within the logic blocks are consistent with the Semi-Modularity property.  Assign the wire values within the logic block to corresponding wires in the half duplicated circuit.  Then assign the value $w$ to corresponding wires in the half duplicated circuit so that $w=c1^A=c2^A=c3^A=w^A$. The same assignment of wire values can also be applied for the B half of a duplicated circuit.
\end{definition}
%\textbf{
For all subsequent discussion in this document, when we talk about the trace of a duplicated circuit we assume that both sides start from the same valid initial state.  I will not include this assumption in the following lemma definitions to avoid complicated wording.  %}  
However for Lemma 1 we only require one half to start from a valid initial state.  
One additional assumption we make about the gates and wire values (including in the presence of faults) are that wire values are strictly 0 or 1, and transitions from excited gates also only produce 0 or 1 and no in between values, no matter the duration of signals from the incoming wires.\\

\begin{lemma}[1]
Let $\sigma_n$ be the sequence of the first n states of a trace $\sigma$ in the duplicated circuit (the trace is possibly finite).  The duplicated circuit has two half circuits $A$ or $B$, if a trace $\sigma$ starts from a valid initial state in the half circuit A, then $h^{A}(\sigma_n)$ is a prefix of a trace of the original circuit.  Also if a trace starts from a valid initial state in the half circuit B, then $h^{B}(\sigma_n)$ is a prefix of a trace of the original circuit.

@@

%If the duplicated circuit is in state $s_n$ and the original circuit is in state $h^A(s_n)$, and for all wires $w$ that are excited in this state $f_w(s_n|_I)\neq w$ in original circuit, then $f_w(s_n|_I) \neq x^A$ or $x^A \neq w^A$ in duplicated circuit (exactly one of these is true).  
%In addition, if wire $w$ is not excited in the original circuit $f_w(s_n|_I)=w$, then in the duplicated circuit half $f_w(s_n|_I)=x^A=w^A$.  
\end{lemma}
\begin{proof}
Without loss of generality we prove that this property holds for the A half of the circuit, the same reasoning can then be applied to the B half of the circuit.  We prove this lemma by induction on all traces of the circuit of length $n$.  The induction hypothesis is
\begin{itemize}
\item For a length $n$ sequence $\sigma_n$ in the duplicated circuit, this maps to $h^{A}(\sigma_n)$, a prefix of a trace of the original circuit.
\item Every logic block is in a Semi-Modularity consistent state.  %newly added, a new property outside of \sigma_n mapping to h^A(\sigma_n)
\item In every state $s_k$ of $\sigma_n$ it is true that a wire $w$ in the original circuit of state $h^A(s_k)$ is excited if and only if exactly one of the following is true: $c1^A$ is excited or $c1^{A}\neq c2^{A}$ or $c2^{A}\neq c3^A$ or $c3^{A}\neq w^A$ (the new value is propagating down one of the logic block or chain of C-elements).  The case where two or more of the three conditions being true does not exist in this setting.  As a consequence a wire $w$ is unexcited if and only if none of the previous three conditions are true. \\
Proving any two of the above three statements is equivalent to the third being true as well.% also I'm calling h^A(s_k) to mean the unduplicated circuit with these states assigned.... forget if I explained or not
\end{itemize}
% true but needs to find a new home -> either $h^A(s_n)=h^A(s_{n+1})$ or $T(h^A(s_n), h^A(s_{n+1}))$
We show in the base case that the induction hypotheses holds for any valid initial state (ie. all possible traces of length 1). 
\begin{itemize}
\item For every $s_1$ that is a valid initial state we have that $h^A(s_1)$ is some state in the original circuit from its construction.  Thus $h^A(s_1)$ is a prefix of size 1 of a trace in the original circuit.  
\item We assigned the wire values so that the logic blocks are in a Semi-Modularity consistent state.  
\item For any wire $w$, since the input wire values are identical in the original circuit and the duplicated half and $c1^A=w$ then $c1^A$ is excited if and only if $w$ is excited in the original circuit (the logic block output is excited).  Together with the definition of the valid initial state $c1^A=c2^A=c3^A=w^A$ (no excitations within the C-element chain), if $w$ is excited in $h^A(s_k)$, then only $c1^A$ is excited.  Conversely if $c1^A$ is excited and $c1^A=c2^A=c3^A=w^A$ then $w$ is excited in $h^A(s_k)$.  All C-element chains in the duplicated have at most one wire in the chain excited. This shows the initial state also satisfies the last part of the induction hypothesis.    %is this enough.... that the existing states follow that iff statement?
\end{itemize}
